---
title: "Air Quality"
author: "Dom Owens"
date: "26/10/2020"
output: html_document
---


```{r}
airqual <- read.csv("air-quality-data-2010-2021.csv") #("air-quality-parson-street.csv")
#stpauls <- read.csv("air-quality-stpauls.csv")
#traffic <- read.csv("fact-traffic-counts.csv")
weather <- read.csv("met-data-bristol-2010-2021.csv") #"met-data-bristol-lulsgate.csv")
head(airqual)
#head(stpauls)
#head(traffic)
head(weather)

```
process airqual
```{r}
library(tidyverse)
airqual <- airqual%>% select(Date.Time,NOx,SiteID)
airqual$Date.Time <- lubridate::as_datetime(airqual$Date.Time)
head(airqual)
```

```{r}
#airqualmap <- read.csv("air-quality-around-bristol.csv")
#head(airqualmap)
unique(airqual$SiteID)
airqual203 <- airqual %>%
 filter(SiteID == 203) %>% select(Date.Time, NOx)
airqual215 <- airqual %>%
 filter(SiteID == 215) %>% select(Date.Time, NOx)
airqual270 <- airqual %>%
 filter(SiteID == 270) %>% select(Date.Time, NOx)
airqual452 <- airqual %>%
 filter(SiteID == 452) %>% select(Date.Time, NOx)
airqual463 <- airqual %>%
 filter(SiteID == 463) %>% select(Date.Time, NOx) ##was 501
dfmap <-airqual203
for(i in list(airqual215,airqual270,airqual452,airqual463)){
  dfmap <- inner_join(dfmap, i, by = "Date.Time", copy=T)
}
dfmap <- dfmap %>% arrange(Date.Time) # in order

#dfmap <- dfmap[157:nrow(dfmap),]## first 156 readings are missing
colnames(dfmap)[2:6] <- c("NOx215","NOx270","NOx452","NOx463","NOx203") 
dfmap[,2:6][which(dfmap[,2:6] ==0, arr.ind = T)] <- 0.01 #prevent -Inf
#dfmap <- na.omit(dfmap)
dfmap[,2:6] <- sqrt(dfmap[,2:6])#log(dfmap[,2:6]) ##take logs
head(dfmap)
```
daily average
```{r}
library("lubridate")
dfdaily <- dfmap %>%
  mutate(day = floor_date(Date.Time, "day")) %>%
           group_by(day) %>%
           summarise_at(2:6,mean, na.rm=T)
head(dfdaily)
plot.ts(dfdaily[,2:6])
```

How to deal with wind direction? Include interaction with speed?
```{r}
plot(dfmap_meteo$Wind.Direction, dfmap_meteo$NOx215)
plot(dfmap_meteo$Wind.Direction, dfmap_meteo$NOx270)#
plot(dfmap_meteo$Wind.Direction, dfmap_meteo$NOx203)
plot(dfmap_meteo$Wind.Direction, dfmap_meteo$NOx452)
plot(dfmap_meteo$Wind.Direction, dfmap_meteo$NOx463)
```

```{r}
library(spatstat)
wd <- dfmap_meteo$Wind.Direction
windrose <- rose(wd,unit="degree", clockwise = T, start = 180, main = "Wind Direction")
```
consider binning into SW/not?


inspect weekend/day effect
```{r}
plot(wday(dfmap_meteo$day, label = T), dfmap_meteo$NOx463)
```


process weather, deal with direction(?), add weekend indicator? or weekday factor?
```{r}
weather$Date.Time <- lubridate::as_datetime(weather$Date.Time) #+ 10 *60 ##align
#weather_trim <- weather %>%
# filter(minute(Date.Time) == 0) ##select on the hour
weatherdaily <- weather %>%
  mutate(day = floor_date(Date.Time, "day")) %>%
           group_by(day) %>%
           summarise_at(2:6,mean, na.rm=T)##daily mean
Max.Wind.Speed<- weather %>%
            mutate(day = floor_date(Date.Time, "day")) %>%
           group_by(day) %>%
           summarise_at(3,max, na.rm=T) ##max windspeed

weatherdaily <- weatherdaily %>% mutate(Max.Wind.Speed = Max.Wind.Speed$Wind.Speed)
colnames(weatherdaily)[3] <- "Mean.Wind.Speed"

#weatherdaily <- weatherdaily %>% 
#  mutate(SinWindDirection = sin(Wind.Direction),CosWindDirection = cos(Wind.Direction))

#weatherdaily <- weatherdaily %>% mutate(Weekend = (wday(day) == 1 |wday(day) == 7 ) ) ##weekend indicator
weatherdaily <- weatherdaily %>% mutate(Weekday = as.factor(wday(day)) )

season <- function(xmonth){
  case_when(
    xmonth %in% c(12,1,2) ~ "Winter",
    xmonth %in% 3:5 ~ "Spring",
    xmonth %in% 6:8 ~ "Summer",
    xmonth %in% 9:11 ~ "Autumn"
  )
}
#season(month(weatherdaily$day[1:5]))
weatherdaily <- weatherdaily %>% mutate(Season = as.factor( season(month(day))) )

head(weatherdaily)
plot.ts(weatherdaily[,2:7]) #take logs of these?
```



Incorporate meteorological data as exogenous regressors
```{r}
dfmap_meteo <- inner_join(dfdaily, weatherdaily, by="day")   ##add meteo data
head(dfmap_meteo)
#logdfmap <- dfmap_meteo %>% select( -c(Wind.Chill.Temperature, Wind.Chill.Temperature.1, Visibility, Dewpoint.Temperature) )
#logdfmap[,2:6] <- log(dfmap_meteo[,2:6]) 
#head(logdfmap)
```

```{r}
colSums(is.na(dfmap_meteo))
#library(zoo)
#logdfmap$Wind.Direction <- na.approx(logdfmap$Wind.Direction) ##interpolate wind direction
dfmap_meteo<- na.omit(dfmap_meteo)
plot.ts(dfmap_meteo[,2:6], main =" " ) #"Hourly NOx Meter Readings (μg/m3), October 2019 - October 2020, Bristol"
#logdfmap$NOx501[which(is.na(logdfmap$NOx501))] <- 0
colSums(is.na( data.frame(dfmap_meteo) ))
any(is.na( data.frame(dfmap_meteo) ))
which(dfmap_meteo < -1000)

```



```{r}
#traffic$Date.Time <- lubridate::as_datetime(traffic$Date.Time)
#stpauls$Date.Time <- lubridate::as_datetime(stpauls$Date.Time)
#unique(traffic$Count.Device.ID) ##some locations have two devices

# traffic16 <- traffic %>%
#  filter(Count.Device.ID == 16) %>% select(Date.Time, Hourly.Flow)
# traffic454 <- traffic %>%
#  filter(Count.Device.ID == 454)%>% select(Date.Time, Hourly.Flow)
# traffic494 <- traffic %>%
#  filter(Count.Device.ID == 494)%>% select(Date.Time, Hourly.Flow)
# traffic380 <- traffic %>%
#  filter(Count.Device.ID == 380)%>% select(Date.Time, Hourly.Flow)
# traffic566 <- traffic %>%
#  filter(Count.Device.ID == 566)%>% select(Date.Time, Hourly.Flow)
# devices <- list(traffic16,traffic454,traffic494,traffic380,traffic566)
# traffic_names <- c("traffic16","traffic454","traffic494","traffic380","traffic566")
```

```{r}
# df <- inner_join(weather_trim, airqual, by="Date.Time")
# df <- inner_join(df, stpauls, by = "Date.Time")
# for(i in 1:length(devices)){
#   df <- inner_join(df, devices[[i]], by = "Date.Time", copy=T)
# }
# colnames(df)[11:15] <- traffic_names
# head(df) ## these are in descending time order
# df <- df %>% arrange(Date.Time)
# head(df)
```




build regression model

```{r}
# airqualmodel <- lm(NOx ~ ., df_na[,c(2:6,10:15)])
# summary(airqualmodel)
# ## do some lasso-based model selection here?
# library(glmnet)
# 
# 
# airqual_lasso_cv <-cv.glmnet(y= df_na$NOx, x= as.matrix(df_na[,c(2:5,10:15)]) )
# airqual_lasso <- glmnet(y= df_na$NOx, x= as.matrix(df_na[,c(2:5,10:15)]), lambda = airqual_lasso_cv$lambda.1se)
# print(airqual_lasso)
# coef(airqual_lasso)
```



```{r}
# 
# df_model <- df_na %>% select("NOx",rownames(coef(airqual_lasso))[c(2:11)] ) #select("NOx","Temperature", "Wind.Speed", "Atmospheric.Pressure","Relative.Humidity", "traffic16","traffic454","traffic494","traffic380")
# library(tidyr)
# #df_model <- coalesce(df_model, (0) )
# sum(is.na(df_model))
# #df_model[is.na(df_model)] <- 0
# #sum(is.na(df_model))
# plot.ts(df_model, main = "Air Quality")

```

### regression models

compare marginal plot for wind direction
```{r}
winddir_model <- lm(NOx215  ~ Temperature + Mean.Wind.Speed +Atmospheric.Pressure + Relative.Humidity + Weekday +Season, #SinWindDirection + CosWindDirection
    data = as.data.frame(dfmap_meteo) )
plot(dfmap_meteo$Wind.Direction,winddir_model$residuals); abline(v = c(180,290), col="red")
```
for season
```{r}
season_model <- lm(NOx215  ~ Temperature + Mean.Wind.Speed +Atmospheric.Pressure + Relative.Humidity + Weekday + Wind.Direction, #SinWindDirection + CosWindDirection
    data = as.data.frame(dfmap_meteo) )
plot(dfmap_meteo$Season,season_model$residuals)
```



final model
```{r}
meteo_models <- lm(cbind(NOx215,NOx270,NOx452,NOx463,NOx203)  ~ Temperature + Mean.Wind.Speed  + Wind.Direction +Atmospheric.Pressure + Relative.Humidity + Weekday +Season, #SinWindDirection + CosWindDirection + Max.Wind.Speed
    data = as.data.frame(dfmap_meteo) )#cbind(NOx215,NOx270,NOx452,NOx463,NOx203)

summary(meteo_models)
```






### residual time series

```{r}
plot.ts(meteo_models$residuals, main = " ")
#par(mfrow = c(3,1))
acf(meteo_models$residuals[,1], lag.max = 28, axes=F, main= " " );axis(1,7*(0:4));axis(2); # abline(v = 7*(1:3), col="green" )
acf(meteo_models$residuals[,1], lag.max = 28*4, axes=F, main= " ", xlab = "Lag"); axis(1,28*(0:4));axis(2) ##monthly
#acf(meteo_models$residuals[,1], lag.max = 28*12*4, axes=F, main= " ", xlab = "Lag"); axis(1,365*(0:4));axis(2)  ##yearly seasonality
```

```{r}
ardfmap <- ar.ols(meteo_models$residuals[1:300,], AIC=T, intercept = T) ## this is actually overspecified - try AIC on post-segments
plot(ardfmap$aic) ##choose p=4?
```

```{r}
# library(Rcpp)
# library(RcppParallel)
# library(RcppArmadillo)
# library(Matrix)
# sourceCpp(file = "/home/do16317/Documents/Changepoints/Wald_RcppParallel.cpp")
# sourceCpp(file = "/home/do16317/Documents/Changepoints/Score_Rcpp.cpp")
```

7 days * 40 weeks =  280

```{r}
library(mosumvar)
airqual_mosum_wald <-  mosumvar::mosumvar(meteo_models$residuals, p=2, G= 280, criterion = "eta", nu = 0.2)
airqual_mosum_wald$plot
#mosum_sub(meteo_models$residuals, p=2, G= 7*52, method = "Wald", criterion = "eta", kap = 1/(52*7))
#logdfmap$`dfmap$Date.Time`[airqual_mosum_wald$cps]
dfmap_meteo$day[airqual_mosum_wald$cps]
airqual_mosum_wald$cps
```

```{r}
airqual_mosum_score <- mosumvar::mosumvar(meteo_models$residuals, method = "Score", p=2, G= 7*40, criterion = "eta", nu = 0.2 )
par(mfrow=c(2,1))
airqual_mosum_wald$plot
airqual_mosum_score$plot
#mosum_sub(meteo_models$residuals, p=3, G= 7*52, method = "Score", criterion = "eta")
#logdfmap$`dfmap$Date.Time`[airqual_mosum_score$cps]
```

acf comparison
```{r}
longest_segment <- meteo_models$residuals[airqual_mosum_wald$cps[5]:airqual_mosum_wald$cps[6],1]
acf(longest_segment, main = "")
```
```{r}
airqual_mosum_score$cps
# par(mfrow = c(4,2), mar = 2*c(1, 2, 1, 1))
# acf(meteo_models$residuals[1:444,1], main = "", xlab = "", ylab = 1, lag.max = 7)
# acf(meteo_models$residuals[445:905,1] , main = "", xlab = "", ylab = 2, lag.max = 7)
# acf(meteo_models$residuals[906:1333,1], main = "", xlab = "", ylab = 3, lag.max = 7)
# acf(meteo_models$residuals[1334:1716,1], main = "", xlab = "", ylab = 4, lag.max = 7)
# acf(meteo_models$residuals[1717:2046,1], main = "", xlab = "", ylab = 5, lag.max = 7)
# acf(meteo_models$residuals[2047:2295,1], main = "", xlab = "", ylab = 6, lag.max = 7)
# acf(meteo_models$residuals[2296:2972,1], main = "", xlab = "", ylab = 7, lag.max = 7)
# acf(meteo_models$residuals[2973:nrow(meteo_models$residuals),1], main = "", xlab = "", ylab = 8, lag.max = 7)
par(mfrow = c(5,1), mar = 2*c(1, 2, 1, 1))
acf(meteo_models$residuals[1:340,1], ylab = "NOx215",main = "", xlab = "",  lag.max = 14)
acf(meteo_models$residuals[1:340,1], ylab = "NOx270",main = "", xlab = "",  lag.max = 14)
acf(meteo_models$residuals[1:340,1], ylab = "NOx452",main = "", xlab = "",  lag.max = 14)
acf(meteo_models$residuals[1:340,1], ylab = "NOx463",main = "", xlab = "",  lag.max = 14)
acf(meteo_models$residuals[1:340,1], ylab = "NOx203",main = "", xlab = "",  lag.max = 14)
```
### model diagnostics
```{r}
norm( ar.ols(meteo_models$residuals, order.max = 2, aic = F)$var.pred, "F" )
norm( ar.ols(meteo_models$residuals[1:370,], order.max = 2, aic = F)$var.pred, "F" )
norm(  ar.ols(meteo_models$residuals[371:881,], order.max = 2, aic = F)$var.pred, "F" )
norm(  ar.ols(meteo_models$residuals[882:1187,], order.max = 2, aic = F)$var.pred, "F" )
norm(  ar.ols(meteo_models$residuals[1188:1513,], order.max = 2, aic = F)$var.pred, "F" )
norm(  ar.ols(meteo_models$residuals[1514:2042,], order.max = 2, aic = F)$var.pred, "F" )
norm(  ar.ols(meteo_models$residuals[2043:2408,], order.max = 2, aic = F)$var.pred, "F" )
norm(  ar.ols(meteo_models$residuals[2409:2703,], order.max = 2, aic = F)$var.pred, "F" )
norm(  ar.ols(meteo_models$residuals[2704:3164,], order.max = 2, aic = F)$var.pred, "F" )


acf(meteo_models$residuals[1:370,1])
acf(meteo_models$residuals[371:881,1])
acf(meteo_models$residuals[882:1187,1])
acf(meteo_models$residuals[1188:1513,1])
acf(meteo_models$residuals[1514:2042,1])
acf(meteo_models$residuals[2043:2408,1])
acf(meteo_models$residuals[2409:2703,1])
acf(meteo_models$residuals[2704:3164,1])
```

```{r}
#compare sSIC
fit_out_model(meteo_models$residuals, cps=c(), pen = log(nrow(x)))$sSIC
fit_out_model(meteo_models$residuals, cps=airqual_mosum_wald$cps, pen = log(nrow(x)))$sSIC
fit_out_model(meteo_models$residuals, cps=airqual_mosum_score$cps, pen = log(nrow(x)))$sSIC
#[1] 11545.51
#[1] 8243.216
#[1] 7894.406
```
## blog

```{r}
gg_df <- data.frame(day =dfmap_meteo$day, meteo_models$residuals)
head(gg_df)

my.ts.panel <- function(x, col = col, bg = bg, pch = pch, type = type,  vpos=airqual_mosum_score$cps, ...){
      lines(x, col = col, bg = bg, pch = pch, type = type, ...)
      abline(v=vpos, col="red")
}

ax <- gg_df$day[seq(1, length(gg_df$day), 350)]

plot.ts(gg_df[,2:6], panel=my.ts.panel, nc =1, xaxt = "n", main = " ", xlab = " " );
axis(1, at = seq(1, length(gg_df$day), 350), labels = ax, las = 1, cex.axis = .5, line = 1.1)#pos = -5.4, outer =F)#axis(1, at = seq(1, length(gg_df$day), 60), labels = ax, las = 2, cex.axis = .5, pos = -3)
# plot(gg_df$NOx215)
# plot(gg_df$NOx270)
# plot(gg_df$NOx452)
# plot(gg_df$NOx463)
# plot(gg_df$NOx203)

# ggplot(gg_df, aes(x=day)) + geom_line(aes(y=NOx215), color="green") + geom_line(aes(y=NOx270),color="blue") + geom_line(aes(y=NOx452),color="grey") + geom_line(aes(y=NOx463),color="purple") + geom_line(aes(y=NOx203),color="orange")
```

